# Red Wine Exploration by Julie Tang

========================================================


```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that will be used in the analysis in this code
# chunk.

library(ggplot2)
library(dplyr)
library(gridExtra)
library(RColorBrewer)
library(scales)
library(GGally)
library(stats)
library(memisc)

```

```{r echo=FALSE, message=FALSE, warning=FALSE, Load_the_Data}
# Load the Data

red_wine <- read.csv('wineQualityReds.csv')

```

This report explores a dataset containing the quality and 
attributes of 1599 red wine samples.


# Univariate Plots Section

Check the structure of the dataset

```{r echo=FALSE, message=FALSE, warning=FALSE, dataset_structure}
str(red_wine)

```

Check if there is any missing data in this dataset

```{r echo=FALSE, message=FALSE, warning=FALSE, any_missing_data}

sapply(red_wine, function(x) sum(is.na(x)))

```

Take a look at the summary data

```{r echo=FALSE, echo=FALSE, message=FALSE, warning=FALSE, dataset_summary}
summary(red_wine)

```


Our dataset consists of 13 variables and 1599 samples.The data type of every
attribute is good and there is no missing data in this dataset. The output
variable, quality, is our interest to analyze.


```{r echo=FALSE, echo=FALSE, message=FALSE, warning=FALSE, Univariate_quality}
      
ggplot(data = red_wine, aes(x = factor(quality))) +
  geom_bar() +
  xlab('quality') +
  ylab('count') 

```
```{r echo=FALSE, echo=FALSE, message=FALSE, warning=FALSE, summary_quality}

summary(red_wine$quality)
table(red_wine$quality)

```


Quality is a discrete quantitative variable. The range of quality variable 
is from 3(lowest) to 8(highest). And the most common level of qualities are 5 and 6. 681 samples'
quality rate is 5 and 638 samples' quality rate is 6.


```{r echo=FALSE, echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_pH}

qplot(data = red_wine, x = pH, binwidth = 0.1, ylab = 'count') +
  scale_x_continuous(breaks = seq(0, 4.01, 0.1))
  
```
```{r, echo=FALSE, message=FALSE, warning=FALSE, summary_pH}

summary(red_wine$pH)

```

The pH is normal distributed. The mean is 3.311 and the median is 3.310, 
which are very close. According to some experts' opinion, pH is one
of the most important attributes on the wine's quality. It affects almost every
aspect of the wine, including the quality. I wondered what's its correlation
with quality in our dataset.


```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_alcohol}

qplot(data = red_wine, x = alcohol, binwidth = 0.2, 
      xlab = 'alcohol(%)', ylab = 'count') + 
  scale_x_continuous(limits = c(8, 15), breaks = seq(8, 15, 0.5))

```

Try log10 transformation since the alcohol distribution is right skewed

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_alcohol_log10}

qplot(data = red_wine, x = alcohol, ylab = 'count', xlab = 'alcohol(log10)(%)') + 
  scale_x_log10()

```

Try square root transformation to see if it will be more close to normal distribution.

```{r echo=FALSE, message=FALSE, warning=FALSE, summary_alcohol}

qplot(data = red_wine, x = alcohol, ylab = 'count', xlab = 'alcohol(sqrt)(%)') + 
  scale_x_sqrt()

```

```{r echo=FALSE, message=FALSE, warning=FALSE, alcohol_summary}

summary(red_wine$alcohol)

```


The distribution of percent alcohol content is right skewed. I tried to use 
log and square root transformation but the distribution was still right skewed. 
It seems the transformation doesn't affect it much. From the plot and summary 
data, we can see the median value of percent alcohol content is about 10. 


```{r echo=FALSE, message=FALSE, warning=FALSE, univariate_plots_sulphates}

qplot(data = red_wine, x = sulphates, 
      xlab = 'sulphates(g / dm^3)', ylab = 'count')

```

```{r echo=FALSE, message=FALSE, warning=FALSE, univariate_plots_sulphates_log}

qplot(data = red_wine, x = sulphates, 
      ylab = 'count', xlab = 'sulphates(log10)(g / dm^3)') + 
  scale_x_log10()

```
```{r echo=FALSE, message=FALSE, warning=FALSE, summary_sulphates}

summary(red_wine$sulphates)

```


The distribution of sulphates is also right skewed. With log10 transformation,
its distribution is normal. From the plot we can see most of samples have
sulphates under 1.0. But there are some outliers with value much higher.


```{r echo=FALSE, message=FALSE, warning=FALSE, univariate_plots_total.so2}

qplot(data = red_wine, x = total.sulfur.dioxide, 
      xlab = 'total.sulfur.dioxide(mg / dm^3)', ylab = 'count')

```

```{r echo=FALSE, message=FALSE, warning=FALSE, univariate_plots_total.so2_log10}

qplot(data = red_wine, x = total.sulfur.dioxide,
      xlab = 'total.sulfur.dioxide(log10)(mg / dm^3)', ylab = 'count') + 
  scale_x_log10()

```
```{r echo=FALSE, message=FALSE, warning=FALSE, summary_total.sulfur.dioxide}

summary(red_wine$total.sulfur.dioxide)

```


The distribution of total sulfur dioxide is also right skewed. With log10
transformation, its distribution is close to normal. From the plot we can see 
most of samples have total sulfur dioxide under 100. But there are a few 
outliers with value around 300!


```{r echo=FALSE, message=FALSE, warning=FALSE, univariate_plots_free.so2}

qplot(data = red_wine, x = free.sulfur.dioxide, 
      xlab = 'free.sulfur.dioxide(mg / dm^3)', ylab = 'count')

```
```{r echo=FALSE, message=FALSE, warning=FALSE, univariate_plots_free.so2_log10}

qplot(data = red_wine, x = free.sulfur.dioxide,
      xlab = 'free.sulfur.dioxide(log10)(mg / dm^3)', ylab = 'count') + 
  scale_x_log10()

```

```{r echo=FALSE, message=FALSE, warning=FALSE, summary_free.sulfur.dioxide}

summary(red_wine$free.sulfur.dioxide)

```


The distribution of free sulfur dioxide has multiple peaks after log10
transformation. One is around 7, the second is around 14 and the third is
around 30.


Create a new variable (all.sulphates) by adding up total.sulfur.dioxide and 
sulphates * 1000.
```{r echo=FALSE, message=FALSE, warning=FALSE, new_variable_all_sulphates}

red_wine <- transform(red_wine, 
                      all.sulphates = total.sulfur.dioxide + sulphates * 1000)

```

```{r, echo=FALSE, message=FALSE, warning=FALSE, univariate_plots_all_sulphates}

qplot(data = red_wine, x = all.sulphates, 
      xlab = 'all.sulphates(mg / dm^3)', ylab = 'count')

```
```{r echo=FALSE, message=FALSE, warning=FALSE, univariate_plots_all_sulphates_log10}

qplot(data = red_wine, x = all.sulphates, 
      xlab = 'all.sulphates(log10)(mg / dm^3)', ylab = 'count') + 
  scale_x_log10()

```

```{r echo=FALSE, message=FALSE, warning=FALSE, summary_all_sulphates}

summary(red_wine$all.sulphates)

```


Since sulfur dioxide is just another type of sulphates. I think maybe we can
add them up to create a new variable, all.sulphates, and check the correlation
of this new variable with quality.

The distribution of Sulphates, total.sulfur.dioxide, free.sulfur.dioxide and
all.sulphates are all right skewed. With the log transformation, they are close
to normal distribution. There are some outliers with very high 
free.sulfur.dioxide and total.sulfur.dioxide. A new variable (all.surphates) is
created to store the total sulphates content in the sample.


```{r echo=FALSE, message=FALSE, warning=FALSE, univariate_plots_fixed.acidity}

qplot(data = red_wine, x = fixed.acidity, 
      xlab = 'fixed.acidity(g / dm^3)', ylab = 'count') +
  scale_x_continuous(limits = c(4, 16), breaks = seq(4, 16, 0.5) )

```

```{r echo=FALSE, message=FALSE, warning=FALSE, summary_fixed.acidity}

summary(red_wine$fixed.acidity)

```

```{r echo=FALSE, message=FALSE, warning=FALSE, univariate_plots_volatile.acidity}

qplot(data = red_wine, x = volatile.acidity, 
      xlab = 'volatile.acidity(g / dm^3)', ylab = 'count')

```


```{r echo=FALSE, message=FALSE, warning=FALSE, summary_volatile.acidity}

summary(red_wine$volatile.acidity)

```


The distribution of both fixed.acidity and volatile.acidity are close to 
normal distribution. But both have some outstanding outliers on upper side.


```{r echo=FALSE, message=FALSE, warning=FALSE, univariate_density}

qplot(data = red_wine, x = density, 
      xlab = 'density(g / dm^3)', ylab = 'count')

```

```{r echo=FALSE, message=FALSE, warning=FALSE, summary_density}

summary(red_wine$density)

```

The distribution of density is normal. And the range is very very small 
(from 0.9901 to 1.0037). It seems it doesn't impact the quality much. 


```{r echo=FALSE, message=FALSE, warning=FALSE, univariate_residual.sugar}

qplot(data = red_wine, x = residual.sugar, 
      xlab = 'residual.sugar(g / dm^3)', ylab = 'count')

```

```{r echo=FALSE, message=FALSE, warning=FALSE, univariate_residual.sugar_log10}

qplot(data = red_wine, x = residual.sugar, 
      xlab = 'residual.sugar(log10)(g / dm^3)', ylab = 'count') + 
  scale_x_log10()

```


```{r echo=FALSE, message=FALSE, warning=FALSE, summary_residual.sugar}

summary(red_wine$residual.sugar)

```


The distribution of residual sugar is right skewed. With the log 
transformation, it's better but still kind of right skewed. And the most of 
sample's sugar level is between 2 and 3.



# Univariate Analysis


### What is the structure of your dataset?

There are 1599 samples in the dataset with 13 variables. 11 are attributes that
might impact the quality (fixed.acidity, volatile.acidity,, citric.acid, 
residual.sugar, chlorides, free.sulfur.dioxide, total.sulfur.dioxide, density, 
pH, sulphates and alcohol). All these attributes are quantitative variables. 
Variable quality is the output variable, which is also the one we are intersted
in. Quality is also quantified.

Other observations:\
1. Most of samples' quality rate is 5 or 6.\
2. The density of all samples are very close. It has a very narrow range(from
0.9901 to 1.0037). It seems it doesn't impact the quality much. \
3. The median percent alcohol content is 10.2. But there are a few outlier 
with value higher than 14.\
4. The median pH value is 3.31 and it's normal distributed.\
5. There are some outliers with very high free.sulfur.dioxide and 
total.sulfur.dioxide. Since high sulphates is not good for health and according 
to experts' opinion, it impact the taste of the wine too. Does higher wine 
quality generally have lower sulphate/Sulfur dioxide content? 


### What is/are the main feature(s) of interest in your dataset?

The main feature of interest in the dataset is quality. I want
to check which attributes have most important impact on the wine's quality.

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

Acidity, residual sugar, chlorides, sulfur dioxide, density, pH, sulphates and 
alcohol likely all contribute to the quality of red wine. I think pH and 
sulphates might contribute most to the quality according to some experts'
opinion on the wine quality.

### Did you create any new variables from existing variables in the dataset?

Yes. I created a new variables named "all.sulphates". Since sulfur dioxide is 
just another type of sulphates, I think it makes sense to add them up together 
to see what impact they have on the quality.

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?

I log-transformed the right skewed free.sulfur.dioxide, total.sulfur.dioxide, 
sulphates and all.surphates distribution. The transformed distribution for these
variables appeal close to normal. The median of all.surphates is 38.56. I also 
noticed there are some outliers with super high surphate content.



# Bivariate Plots Section

Remove the first variable (X) before generating the correlation matrix, since 
X is just an id for every sample
```{r echo=FALSE, message=FALSE, warning=FALSE, pairs_plot}
# remove the first variable (X) before generating the correlation matrix, since 
# X is just an id for every sample.

red_wine_temp <- subset(red_wine, select = -c(X, all.sulphates))

ggcorr(red_wine_temp,
       label = TRUE,
       label_alpha = TRUE,
       label_round = 2,
       layout.exp = 2,
       hjust = 1)

#ggpairs(red_wine_temp) +
#  theme(axis.ticks = element_blank(),
#        axis.text = element_blank())

```


From the correlation matrix above, it looks to me that many attributes tend to 
correlate with each other and also correlate with quality. But I don't see 
either two have very strong correlation. I think it may be because the wine 
quality somehow is very difficult to accurately describe and quantify. I suppose
pH and sulphates impact the quality most according to the articles I read from 
the website. However, from this plot, it looks alcohol has most correlation with
the quality, very interesting.



```{r echo=FALSE, message=FALSE, warning=FALSE, summary_pH_quality_Plots}

by(red_wine$pH, red_wine$quality, summary)

```


```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Scatter_Plots_pH}

ggplot(aes(x = pH, y = quality),
       data = red_wine) +
  geom_jitter(alpha = 0.25) +
  stat_summary(fun.y = mean, geom = 'line', color = 'red')
  

```

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_correlatio_pH_quality}

cor.test(red_wine$pH, red_wine$quality)

```


From the scatter plot above, I can't tell if there is any correlation between
pH and quailty. The Peason's correlation test also shows there is very week 
correlation between them. The correlation value is just -0.058. Surprised!


```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots_sulphates}

ggplot(aes(x = sulphates, y = quality),
       data = red_wine) +
  geom_jitter(alpha = 0.25) +
  xlab('sulphates(g / dm^3)')
  

```

Do a log10 transformation and only check the 95% of sulphates data to see if
the shape will be better.
```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots_sulphates_log10}

ggplot(aes(x = sulphates, y = quality), data = red_wine) +
  geom_jitter(alpha = 0.25) +
  stat_summary(fun.y = mean, geom = 'line', color = 'red') +
  scale_x_log10() +
  xlab('sulphates(log10)(g / dm^3)') +
  coord_cartesian(xlim = c(0.3, quantile(red_wine$sulphates, 0.95)))
  

```


```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_summary_sulphates_quality}

by(red_wine$sulphates, red_wine$quality, summary)

```

```{r echo=FALSE, message=FALSE, warning=FALSE, correlation_sulphates_quality}

cor.test(red_wine$sulphates, red_wine$quality)

```


It's very hard to tell if there is any relationship between the sulphates and
quality from the first scatter plot. By log10 transformationSince most samples
and look closer to the data(95%), we can see the mean quality is a slightly 
trending up along with sulphate value. This is also surprising since I suppose
higher quality has lower sulphate content. The correlation value is 0.251, which
indicates a weak correlation.


```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots_all_sulphates}

ggplot(aes(x = all.sulphates, y = quality), binwidth = 30, data = red_wine) +
  geom_jitter() +
  scale_x_log10() +
  xlim(min(red_wine$all.sulphates), quantile(red_wine$all.sulphates, 0.99)) +
  xlab('all.sulphates(mg / dm^3)') +
  geom_smooth() 
  

```

```{r echo=FALSE, message=FALSE, warning=FALSE, correlation_all_sulphates_quality}

cor.test(red_wine$all.sulphates, red_wine$quality)

```

It seems it's not good idea to combine sulphates and sulfur dioxide. With the
combined sulphates I don't see there is any obvious correlation between 
sulphates and quality, also the correlation value decreased from 0.251 to 0.21. 
Maybe we should look sulphates and sulfur dioxide data separately.


```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_boxPlots_sulfur_dioxide}

ggplot(aes(x = total.sulfur.dioxide, y = quality),
       data = red_wine) +
  geom_jitter() +
  scale_x_log10() +
  stat_summary(fun.y = mean, geom = 'line', color = 'red') +
  xlab('total.sulfur.dioxide(mg / dm^3)')
  

```

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_boxPlots_free_so2}

ggplot(aes(x = free.sulfur.dioxide, y = quality),
       data = red_wine) +
  geom_jitter() +
  scale_x_log10() +
  stat_summary(fun.y = mean, geom = 'line', color = 'red') +
  xlab('free.sulfur.dioxide(mg / dm^3)')
  

```


yes, it looks either free sulfur dioxide or total sulfur dioxide content impact
the quality much.  But Sulphates content does look be correlated, though not 
strong, with the quality level. 


From the pairs plot, I noticed there are some other interesting relationships
between the attributes. For example fixed.acidity and pH, fixed.acidity and 
density, fixed.acidity and citric.acid, free.sulfur.dioxide and 
total.sulfur.dioxide. 


```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_plots_acidity_pH}

ggplot(aes(x = pH, y = fixed.acidity), data = red_wine) +
  geom_point(alpha = 0.25) +
  geom_smooth() +
  ylab('fixed.acidity(g / dm^3)')

```

```{r echo=FALSE, message=FALSE, warning=FALSE, Correlation_pH_acidity}

cor.test(red_wine$pH, red_wine$fixed.acidity)

```


Both the scatter plot and the Pearson's correlation test show there are 
negative correlation between the pH and fixed.acidity. The higher pH the lower
acidity. And the correlation is strong ( -0.683).


```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_plots_free_total_s02}

ggplot(aes(x = free.sulfur.dioxide, y = total.sulfur.dioxide),
       data = red_wine) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = 'lm') +
  xlab('free.sulfur.dioxide(mg / dm^3)') +
  ylab('total.sulfur.dioxide(mg / dm^3)')


```

```{r echo=FALSE, message=FALSE, warning=FALSE, Correlation_free_total_so2}

cor.test(red_wine$free.sulfur.dioxide, red_wine$total.sulfur.dioxide)


```

```{r echo=FALSE, message=FALSE, warning=FALSE, lm_free_total_so2}
# create a linear model

m1 = lm(total.sulfur.dioxide ~ free.sulfur.dioxide, data = red_wine)
mtable(m1, sdigits = 3)


```


The scatter plot and Pearson's correlation test both show there are 
strong correlation between free sulfur dioxide content and total sulfur dioxide
content. And it's close to a linear model. By creating the linear model however
it can only explain 44.6% of the result. It looks it's not a strong 
linear relationship.


```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_plots_alcohol_quality}

ggplot(aes(x = alcohol, y = quality), data = red_wine) +
  geom_jitter(alpha = 0.25) +
  geom_smooth(method = 'lm')


```

```{r echo=FALSE, message=FALSE, warning=FALSE, Correlation_alcohol_quality}

cor.test(red_wine$alcohol, red_wine$quality)


```


Among all the attributes, alcohol has the strongest correlation (0.476) with 
quality. I'm surprised since Alcohol is never mentioned in the articles I have 
read.


# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset

From the pairs plot it looks all the attributes somewhat affect the quality. I 
choosed to do some further investigation on the three attributes which I think 
are most important, pH, sulphates and Alcohol. From the scatter plots, I 
couldn't tell there is much correlation between pH and quality. But Sulphates 
and Alcohol do have direct correlation with the wine's quality. The higer wine 
quality is associated with higher sulphates. I'm confused on this finding since 
I suppose the wine with higher quality should have less suphates. But the 
range of suphates is very narrow(between 0.3 and 1). The higher quality 
generally associated with higher alcohol content.

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

Yes. There are some interesting relationships between the attributes. 
* There are negative correlation between fixed.acidity and pH, the higher pH 
the lower acidity. 
* Between free.sulfur.dioxide and total.sulfur.dioxide, Generally higher free 
sulfur dioxide higher total sulfur dioxide. This is easy to understand.

### What was the strongest relationship you found?

The strongest relationship I found is between attribute fixed.acidity and pH. 
Its correlation value is -0.6829782.


# Multivariate Plots Section


```{r echo=FALSE, message=FALSE, warning=FALSE, quality.bucket}

red_wine$quality.bucket <- cut(red_wine$quality, breaks = c(2, 4, 6, 8))

table(red_wine$quality.bucket)

```


Quality rates are categorized into 3 buckets: (2,4] low quality; (4,6] Medium
quality; (6, 8] high quality to assist the analysis of  multiple variables.


```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_plots_pH_acidity_quality_bucket}

ggplot(aes(x = pH, y = fixed.acidity), data = red_wine) +
  geom_jitter(aes(color = quality.bucket)) +
  geom_smooth(method = 'lm', aes(color = quality.bucket)) +
  ylab('fixed.acidity(g / dm^3)')
  

```


It shows the pH and fixed acidity have negative correlation in all three 
quality buckets. 


```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_plots_free_total_sulfur_dioxide_quality}

ggplot(aes(x = free.sulfur.dioxide, y = total.sulfur.dioxide),
       data = red_wine) +
  geom_jitter(aes(color = quality.bucket)) +
  geom_smooth(method = 'lm', aes(color = quality.bucket)) +
  xlab('free.sulfur.dioxide(mg / dm^3)') +
  ylab('total.sulfur.dioxide(mg / dm^3)')
  

```


It shows the free sulfur dioxide and total sulfur dioxide have positive 
correlation in all three quality buckets.  


```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_plots_pH_quality_acidity}

ggplot(aes(x = pH, y = quality, color = fixed.acidity),
       data = red_wine) +
  geom_jitter(alpha = 0.5) + 
  scale_color_distiller(type = 'div') +
  ggtitle('Wine quality with pH by fixed aciditity')

```


I can't tell there is much correlation between pH and quality from this 
scatter plot. But after adding the fixed acidity, we can see higher fixed 
acidity is on the left, associated with lower pH. It's becauase fixed acidity 
are negatively correlated with pH.


```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_plots_pH_alcohol_quality}

ggplot(aes(x = alcohol, y = quality, color = pH),
       data = red_wine) +
  geom_jitter(alpha = 0.5) +
  scale_color_distiller(type = 'div') +
  ggtitle('Wine quality with alcohol by pH')
  

```


The correlation between alcohol and quality is supposed to be the strongest
comparing to other attributes. By adding pH, we can see different pH levels 
spread the whole plot. But it looks higher pH cluster more on higher
percent alcohol side. And lower pH culster more on lower percent alcohol side.
The pair plots show pH and Alcohol's col value is 0.206. So they are somewhat 
correlated.


```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_plots_sulphates_quality_pH}

ggplot(aes(x = sulphates, y = quality, color = pH),
       data = red_wine) +
  geom_jitter(alpha = 0.5) +
  scale_x_log10() +
  xlab('sulphates(log10)(mg / dm^3)') +
  scale_color_distiller(type = 'div') +
  ggtitle('Wine quality with sulphates by pH')
  

```


In the bivariate plots we have concluded sulphates and quality are correlated.
In this plot after adding pH, we can see samples with different pH spread 
everywhere. Again it seems pH doesn't have much correlation with quality. From
this plot, it's interesting to see the samples with sulphates higher than 1 
generally have lower pH and associated with quality level lower than 6.


```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_plots_alcohol_sulphates_quality}

ggplot(aes(x = alcohol, y = quality, color = sulphates),
       data = red_wine) +
  geom_jitter(alpha = 0.5) +
  scale_color_distiller(type = 'div') +
  ggtitle('Wine quality with alcohol by sulphates')
  

```


Both alcohol and sulphates have positive correlation with quality.  When 
putting these three variables together, it looks there is a mild trend of higher
alcohol, higer sulphates(under 1) have higher quality. But the correlation is 
not strong. 


```{r echo=FALSE, message=FALSE, warning=FALSE,multivariate_lm}

m1 <- lm(quality ~ alcohol, data = red_wine)
m2 <- update(m1, ~ . + pH)
m3 <- update(m2, ~ . + I(log(sulphates)))
m4 <- update(m3, ~ . + volatile.acidity)
mtable(m1, m2, m3, m4, sdigits = 3)

```


The variables in this linear model can only account to 34.9% of the variance
in the quality of red wine. We continue to do the log transformation on 
Sulphates.

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

To be honest, I don't see any attribute has strong correlation with quality.
It's hard to say which attribute is most critical to predict the wine's quality.
Though from correlation coefficient's perspective, the alcohol is most 
correlated with quality, the corrlation is still not strong.

But when putting alcohol, sulphates and quality in the same plot, we can see 
generally the samples with higher alcohol and higher sulphates(but lower than 1)
has the trend to be higher quality. This is reasonable because both alcohol and
sulphate (under 1) have positive correlation with quality.



### Were there any interesting or surprising interactions between features?
pH seems can't predict the quality at all. While doing the research to 
understand what might contribute to wine's quality, pH seems always be the top 
one. I suppose the plot should show some correlation between pH and quality. But
no matter where I add pH it spreads everywhere. 


### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.
Yes, I created a linear model that included alcohol, sulphates, pH and volatile
acidity variables and wine quality as the output variable. The
limitation of this linear model is that variables in this linear model can only 
account for 34.9% of the variance in the quality of red wine. The strengths is
that we can still tell some correlations between the predictors and output
variable and by adding additonal variable we got higher r-squared value.

------

# Final Plots and Summary


### Plot One
```{r echo=FALSE, echo=FALSE, message=FALSE, warning=FALSE, Plot_One}

ggplot(aes(x = pH, y = quality, binwidth = 0.1),
       data = red_wine) +
  geom_jitter(alpha = 0.25) +
  scale_x_continuous(limits = c(quantile(red_wine$pH, 0.01), 
                                quantile(red_wine$pH, 0.99))) +
  geom_line(stat = 'summary', fun.y = median, color = 'red') +
  geom_smooth() +
  ggtitle('Wine quality by pH')

```


### Description One

This Plot shows the relationship between pH and quality. I excluded the samples
with bottom 1% and top 1% of pH to explain the majority of samples to avoid
the interrupt of extreme cases. Though there is no direct correlation between pH
and quality, we can see pH is mostly between 3.1 and 3.5 for the samples with 
high quality(7 or 8). Since the majority samples' pH is between 3.0 and 3.7, 
the range is narrow. Maybe that can also explain why we don't see much 
correlation between pH and quality. When pH is above 3.6, we can see there are
few samples with high quality. If the pH's range is bigger, maybe we can see its
impact on the quality.

### Plot Two

```{r echo=FALSE, echo=FALSE, message=FALSE, warning=FALSE, Plot_Two}

ggplot(aes(x = pH, y = fixed.acidity),
       data = red_wine) +
  geom_jitter(aes(color = quality.bucket)) +
  geom_smooth(method = 'lm', aes(color = quality.bucket)) +
  ggtitle('Fixed acidity by pH and quality bucket') +
  ylab('fixed acidity(g / dm^3)')

```

### Description Two

The plot above shows the correlation between pH and fixed acidity. The
correlation between these two attributes are the strongest among all the
attributes. I split the quality level into 3 buckets:(2,4],(4, 6], (6, 8]. It's 
interesting to see that if we use lm smoothing method, the slope of higher 
quality is steeper than the lower quality.


### Plot Three

```{r echo=FALSE, echo=FALSE, message=FALSE, warning=FALSE, Plot_Three}

ggplot(aes(x = alcohol, y = quality, color = sulphates),
       data = red_wine) +
  geom_jitter(alpha = 0.5) +
  scale_x_continuous(limits = c(min(red_wine$alcohol), 
                                quantile(red_wine$alcohol, 0.99))) +
  scale_color_distiller(type = 'div') +
  geom_smooth() +
  ggtitle('Wine quality by alcohol and sulphates') +
  xlab('alcohol(%)')

```

### Description Three

Both alcohol and sulphates have positive correlation with quality.  When putting 
these three variables together, we can see there is a mild trend: higher alcohol,
higer sulphates(under 1) have higher quality. But the correlation is not strong.
Also, there are few samples have sulphates higher than 1.0. 

------

# Reflection

When I started to explore the red wine dataset, I was very confused on what each
variable means since I'm not familiar with wine at all. I did some research on 
google and read some articles to get better understanding. According to some 
articles, I believe pH and Sulphates are the most important attributes that 
impact the quality. So I decided to focus on the relationship between these two 
attributes and quality.

However, I was surprised I didn't see strong correlations between them and 
quality at all, especially pH. In the plots sulphates shows mild positive
correlation with quality(corr is 0.251). I had been thinking why this happened 
during the whole exploration. Surprisingly, Alcohol shows the strong correlation
with quality among all the attributes(corr is 0.476). 

After I explore more and more, I think the reason this happened is because that
there is some limitation on the dataset. There are 1599 samples in total. And 
1,319 out of 1,599 received a score of 5 or 6. No sample received scores of 0, 
1, 9, or 10. Since the wine quality variable had such limited variability, it 
was difficult to assess the relationship between quality and the chemical 
attribute variables. If we have a greater variety of quality ratings we might 
get a better insight on how these attributes are related with quality.
Also, quality rating is based on sensory data (median of at least 3 evaluations 
made by wine experts). It's difficult to accurately describe and quantify. 

Besides the attributes in the dataset, there are some other aspects that impact
the quality, like the type of grape, the place and climate where the grape grow
etc. There is another dataset of white wine, which I would like to do another
exploration to compare with the data of red wine to see if there is any different
pattern.



# Reference documents:

http://www.wineskills.co.uk/winemaking/winemaking-knowledge-base/sulphur-dioxide

http://winemakersacademy.com/importance-ph-wine-making/
